### 2019-07-15 draft_charts.Rmd
Change legend in chart 2 and title in chart 7

---
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: hide
    toc: TRUE
    toc_float: TRUE
    css: www/web_report.css
    editor_options:
      chunk_output_type: console
self_contained: no      
editor_options:
  chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

![](www/images/urban-institute-logo.png)

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

knitr::opts_chunk$set(fig.width = 6.5, fig.height = 6, fig.path = "figures/", dev = c("png", "cairo_pdf"))
```

```{r setup}
options(scipen = 999)

library(tidyverse)
library(urbnthemes)
library(grid)
library(gridExtra)
library(extrafont)

source("urban_style.R")

set_urbn_defaults(style = "print")
```

### N Charts About Future Retirement Security plots
### Aaron R. Williams
### `r format(Sys.time(), "%B %d, %Y %H:%M %p")`

```{r}
cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")

#' rename_cohort
#'
#' @param x a vector with cohorts
#'
#' @return a vector of renamed cohorts
rename_cohort <- function(x, two_lines = TRUE) {
  if (two_lines == TRUE) {
    cohort <- recode(x,
      `1936-1945` = "Pre-boomers\n(1936–1945)", 
      `1946-1955` = "Early boomers\n(1946–1955)", 
      `1956-1965` = "Late boomers\n(1956–1965)", 
      `1966-1975` = "Gen Xers\n(1966–1975)", 
      `1976-1985` = "Xennials\n(1976–1985)")

    factor(cohort, levels = c("Pre-boomers\n(1936–1945)", 
                              "Early boomers\n(1946–1955)", 
                              "Late boomers\n(1956–1965)", 
                              "Gen Xers\n(1966–1975)", 
                              "Xennials\n(1976–1985)"))
  } else {
    cohort <- recode(x,
      `1936-1945` = "Pre-boomers (1936–1945)", 
      `1946-1955` = "Early boomers (1946–1955)", 
      `1956-1965` = "Late boomers (1956–1965)", 
      `1966-1975` = "Gen Xers (1966–1975)", 
      `1976-1985` = "Xennials (1976–1985)")

    factor(cohort, levels = c("Pre-boomers (1936–1945)", 
                              "Early boomers (1946–1955)", 
                              "Late boomers (1956–1965)", 
                              "Gen Xers (1966–1975)", 
                              "Xennials (1976–1985)"))
  }
}

#' to_ordinal
#'
#' @param x translate numeric integers into ordinal numbers in a character vector 
#'
#' @return
#' @export
#'
#' @examples to_ordinal(1:30)
to_ordinal <- function(x, capitalize = FALSE, as_factor = FALSE) {
  # case_when(x %in% 11:13 ~ paste0(x, "th"),
  #           x %% 10 == 0 ~ paste0(x, "th"),
  #           x %% 10 == 1 ~ paste0(x, "st"),
  #           x %% 10 == 2 ~ paste0(x, "nd"),
  #           x %% 10 == 3 ~ paste0(x, "rd"),
  #           x %% 10 %in% 4:9 ~ paste0(x, "th"),
  #           TRUE ~ "ERROR")

  word <- case_when(x == 0 ~ "Zeroth",
                    x == 1 ~ "Bottom",
                    x == 2 ~ "Second",
                    x == 3 ~ "Third",
                    x == 4 ~ "Fourth",
                    x == 5 ~ "Top")

  if (capitalize == TRUE) {
    tools::toTitleCase(word)
  } else {
    word
  }
}
```

## Theme 1: Retirement incomes will continue to rise

```{r theme1}
chart2 <- read_csv("data/figure2.csv") %>%
  filter(doby %in% cohorts) %>%
  mutate(doby = rename_cohort(doby)) %>%  
  ggplot(aes(doby, pc_cen_net_income_P50)) +
  geom_col() +
  geom_text(mapping = aes(label = scales::dollar(signif(pc_cen_net_income_P50, 3), largest_with_cents = 0)), 
            vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()

# create the output plot
plotr(titler("A typical Xennial retiree will receive about 30 percent more net income than a typical retiree born 40 years earlier", width = 60),
      subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
      chart2,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("This income measure includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, capital gains, and retirement account withdrawals, minus payroll taxes, federal and state income taxes, and Medicare surtaxes. The analysis computes a per capita measure by dividing total family income by two for married people and assumes that scheduled Social Security benefits are paid."),
      heights = c(4, 1.5, 30, 1, 1, 4))
```

```{r}
data2 <- read_csv("data/figure2.csv") %>%
  filter(doby %in% cohorts)

data2[data2$doby == "1966-1975", "pc_cen_net_income_P50"] / data2[data2$doby == "1936-1945", "pc_cen_net_income_P50"] - 1

data2[data2$doby == "1976-1985", "pc_cen_net_income_P50"] / data2[data2$doby == "1936-1945", "pc_cen_net_income_P50"] - 1

```

## Theme 2: Women’s increased earnings will power retirement income growth

```{r theme2}
chart4 <- read_csv("data/figure4.csv") %>%
  mutate(male = recode(male, 
                       `1.Male` = "Men",
                       `2.Female` = "Women")) %>%
  filter(doby %in% cohorts) %>%
  filter(!is.na(male)) %>%
  mutate(doby = rename_cohort(doby)) %>%  
  ggplot(aes(doby, lifetime_earn_P50, fill = male)) +
  geom_col(position = "dodge") +
  geom_text(mapping = aes(label = scales::dollar(round(lifetime_earn_P50, -2), largest_with_cents = 0)), 
            position = position_dodge(width = 1), vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()

# create the output plot
plotr(titler("Over her lifetime, a typical Xennial woman will earn more than twice as much as her pre-boomer counterpart"),
      subtitler("Median earnings from ages 25 to 69 (constant 2018 dollars)"),
      chart4,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("Estimates are restricted to adults who survive to age 70."), 
      heights = c(4, 2, 30, 1, 1, 1))
```

```{r}
data4 <- read_csv("data/figure4.csv") %>%
  filter(doby %in% cohorts)

data4[data4$doby == "1966-1975" & data4$male == "2.Female", "lifetime_earn_P50"] / data4[data4$doby == "1936-1945" & data4$male == "2.Female", "lifetime_earn_P50"] - 1

data4[data4$doby == "1976-1985" & data4$male == "2.Female", "lifetime_earn_P50"] / data4[data4$doby == "1936-1945" & data4$male == "2.Female", "lifetime_earn_P50"] - 1

data4[data4$doby == "1976-1985" & data4$male == "1.Male", "lifetime_earn_P50"] / data4[data4$doby == "1936-1945" & data4$male == "1.Male", "lifetime_earn_P50"] - 1

data4[data4$doby == "1976-1985" & data4$male == "1.Male", "lifetime_earn_P50"] / data4[data4$doby == "1976-1985" & data4$male == "2.Female", "lifetime_earn_P50"] - 1
```

## Theme 3: Despite growth in retirement incomes, retirement resources will fall short for more Gen Xers and Xennials

* Numerator: age 70, price-adjusted, per-capita annuitized income
* Denominator: mean price-adjusted, per-capita earnings from ages 50 to 59

```{r theme3}
chart1 <- read_csv("data/figure1.csv") %>%
  filter(ann_reprate %in% c("0-50%", "50-75%")) %>%
  filter(doby %in% cohorts) %>%
  mutate(doby = rename_cohort(doby)) %>%
  group_by(doby) %>%
  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
  ggplot(aes(doby, rep_rate_threshold)) +
  geom_col() +
  geom_text(mapping = aes(label = scales::percent_format(accuracy = 1)(rep_rate_threshold / 100)), vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
                     limits = c(0, 40)) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()

# create the output plot
plotr(titler("Nearly one-third of Xennials will see their living standards fall when they retire", width = 61),
      subtitler("Share of adults unable to replace 75 percent of preretirement earnings at age 70"),
      chart1,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("The analysis compares per capita family income at age 70 to average annual family earnings received from ages 50 to 59, both measured in constant 2018 dollars as adjusted by the change in the consumer price index. The income measure at age 70 includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, and the annual income from an annuity valued at 80 percent of financial assets (including retirement accounts). The analysis computes a per capita measure by dividing total family income by two for married people and assumes that scheduled Social Security benefits are paid."), 
      heights = c(4, 2, 30, 1, 1, 5))
```

## Theme 4: Social Security’s financing problems threaten retirement security

```{r, theme4}

scheduled <- read_csv("data/figure1.csv") %>%
  mutate(law = "Current Social Security benefits continue")

payable <- read_csv("data/figure9.csv") %>%
  mutate(law = "Benefits cut 25%") %>%
  rename(ann_reprate = ann_payable_reprate)

chart9 <- bind_rows(scheduled, payable) %>%
  filter(ann_reprate %in% c("0-50%", "50-75%")) %>%
  filter(doby %in% cohorts) %>%
  group_by(doby, law) %>%
  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
  ungroup() %>%
  mutate(law = factor(law, levels = c("Current Social Security benefits continue", "Benefits cut 25%"))) %>%
  mutate(doby = rename_cohort(doby)) %>%  
  ggplot(aes(doby, rep_rate_threshold, fill = law)) +
  geom_col(position = "dodge") +
  geom_text(mapping = aes(label = scales::percent_format(accuracy = 1)(rep_rate_threshold / 100)), 
            vjust = -1,
            position = position_dodge(width = 0.8)) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()

# create the output plot
plotr(titler("Retirement incomes may fall short for about 4 in 10 Xennials if policymakers ignore Social Security’s financing gap"),
      subtitler("Share of adults unable to replace 75 percent of their preretirement earnings at age 70"),
      chart9,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("The analysis compares per capita family income at age 70 to average annual earnings received from ages 50 to 59, both measured in constant 2018 dollars as adjusted by the change in the consumer price index. The income measure at age 70 includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, and the annual income from an annuity valued at 80 percent of financial assets (including retirement accounts). The analysis computes a per capita measure by dividing total family income by two for married people.", width = 131),
      heights = c(6.5, 1.5, 30, 1, 1, 5))
```

## Theme 5: Social Security cuts would disproportionately harm low-income retirees

* Quintiles are based on per-capita, price-adjusted, shared annuitized income with federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70.

```{r theme5}
quintile <- c("Bottom quintile", "Second quintile", "Third quintile", "Fourth quintile", "Top quintile")

chart7alt <- read_csv("data/figure7b.csv") %>%
  mutate(nchartQuin = paste(to_ordinal(nchartQuin), "quintile")) %>%
  mutate(nchartQuin = factor(nchartQuin, levels = quintile)) %>%
  mutate(ss_share = PCssb_Mean / pc_cen_income_Mean) %>%
  filter(doby %in% cohorts) %>%
  mutate(doby = rename_cohort(doby)) %>%    
  select(nchartQuin, doby, ss_share) %>%
  ggplot(aes(doby, ss_share, fill = nchartQuin)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = scales::percent_format(accuracy = 1)(ss_share)),
            position = position_dodge(width = 0.7),
            vjust = -0.8,
            hjust = 0.35) +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
                     labels = scales::percent) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()

# create the output plot
plotr(titler("Social Security accounts for about 85 percent of income for low-income beneficiaries"),
      subtitler("Share of per capita family income at age 70 from Social Security by income quintile"),
      chart7alt,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("This income measure includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, capital gains, and retirement account withdrawals. The analysis computes a per capita measure by dividing total family income by two for married people. The sample is restricted to Social Security beneficiaries. The analysis assumes that scheduled Social Security benefits are paid."),
      heights = c(4, 1, 30, 1, 1, 4))
```

## Theme 6: Old-age poverty rates could rise if policymakers ignore Social Security’s financial problems

```{r, theme6}
chart12 <- read_csv("data/figure12.csv") %>%
  filter(doby %in% cohorts) %>%
  rename(`Current Social Security benefits continue` = poverty_status_Mean, 
         `Benefits cut 25%` = poverty_status_payable_Mean) %>%
  mutate(doby = rename_cohort(doby)) %>% 
  gather(key = "law", value = "poverty", -doby, -Obs) %>%
  mutate(law = factor(law, levels = c("Current Social Security benefits continue", "Benefits cut 25%"))) %>%
  ggplot(aes(doby, poverty, fill = law)) +
  geom_col(position = "dodge") +
  geom_text(mapping = aes(label = scales::percent_format(accuracy = 0.1)(poverty)), 
            vjust = -1,
            position = position_dodge(width = 0.8)) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()

# create the output plot
plotr(titler("A 25 percent Social Security benefit cut could leave about 1 in 12 Gen X and Xennial retirees in poverty"),
      subtitler("Share of adults with incomes below the federal poverty level at age 70 "),
      chart12,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("The analysis compares family income from Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, rent, retirement account withdrawals, and income from coresident family members to the federal poverty level. "),
      heights = c(4, 1.5, 30, 1, 1, 3))
```

## Theme 7: Racial and ethnic disparities in retirement income will persist

```{r theme7}
chart6 <- read_csv("data/figure6.csv") %>%
  filter(doby %in% cohorts) %>%
  filter(!is.na(raceeth)) %>%
  filter(raceeth != "4.Other") %>%
  mutate(raceeth = recode(raceeth, 
                          `1.White, Non-Hisp` = "White, non-Hispanic",
                          `2.Black, Non-Hisp` = "Black, non-Hispanic",
                          `3.Hispanic` = "Hispanic"
                          )) %>%  
  mutate(doby = rename_cohort(doby)) %>%  
  ggplot(aes(doby, pc_cen_net_income_P50, fill = raceeth)) +
  geom_col(position = "dodge") +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15)), 
                     labels = scales::dollar) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks()

# create the output plot
plotr(titler("Although retirement incomes are projected to rise for people of color, they will continue to lag far behind incomes for white people"),
      subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
      chart6,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("This income measure includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, capital gains, and retirement account withdrawals, minus payroll taxes, federal and state income taxes, and Medicare surtaxes. The analysis computes a per capita measure by dividing total family income by two for married people. The analysis assumes that scheduled Social Security benefits are paid. The chart excludes other racial groups because of small sample sizes."),
      heights = c(6.5, 1.2, 28, 1, 1, 4.5))
```

```{r}
data6 <- read_csv("data/figure6.csv") %>%
  filter(doby %in% cohorts)

1 - data6[data6$doby == "1936-1945" & data6$raceeth == "2.Black, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"] / data6[data6$doby == "1936-1945" & data6$raceeth == "1.White, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"]

(data6[data6$doby == "1936-1945" & data6$raceeth == "3.Hispanic" & !is.na(data6$raceeth), "pc_cen_net_income_P50"] - data6[data6$doby == "1936-1945" & data6$raceeth == "1.White, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"]) / data6[data6$doby == "1936-1945" & data6$raceeth == "1.White, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"]
```

## Theme 8: The gender gap in Social Security benefits will narrow

```{r, theme8}
chart3b <- read_csv("data/figure3b.csv") %>%
  mutate(male = recode(male, 
                       `1.Male` = "Men",
                       `2.Female` = "Women")) %>%  
  filter(doby %in% cohorts) %>%
  mutate(doby = rename_cohort(doby)) %>%  
  ggplot(aes(doby, Ownssb_P50, fill = male)) +
  geom_col(position = "dodge") +
  geom_text(mapping = aes(label = scales::dollar(round(Ownssb_P50, -2), largest_with_cents = 0)), 
            position = position_dodge(width = 0.75), vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()

# create the output plot
plotr(titler("For Xennials, women’s median Social Security benefits at age 70 will rise to 85 percent of the median for men"),
      subtitler("Median annual Social Security income at age 70 (constant 2018 dollars)"),
      chart3b,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("Analysis assumes that promised Social Security benefit payments will continue."), 
      heights = c(4, 1, 30, 1, 1, 1))
```

```{r}
data3b <- read_csv("data/figure3b.csv") %>%
  filter(doby %in% cohorts)

data3b[data3b$doby == "1976-1985" & data3b$male == "1.Male", "Ownssb_P50"] / data3b[data3b$doby == "1936-1945" & data3b$male == "1.Male", "Ownssb_P50"]

data3b[data3b$doby == "1976-1985" & data3b$male == "2.Female", "Ownssb_P50"] / data3b[data3b$doby == "1936-1945" & data3b$male == "2.Female", "Ownssb_P50"]

1 - data3b[data3b$doby == "1936-1945" & data3b$male == "2.Female", "Ownssb_P50"] / data3b[data3b$doby == "1936-1945" & data3b$male == "1.Male", "Ownssb_P50"]

1 - data3b[data3b$doby == "1976-1985" & data3b$male == "2.Female", "Ownssb_P50"] / data3b[data3b$doby == "1976-1985" & data3b$male == "1.Male", "Ownssb_P50"]
```

## Theme 9: Gen Xers and Xennials will pay more taxes in retirement

```{r, theme9}

chart13 <- read_csv("data/figure13.csv") %>%
  filter(doby %in% cohorts) %>%
  mutate(avg_annual_tax_rate = (-1 * pc_tot_tax_Mean) / pc_cen_income_Mean) %>%
  mutate(doby = rename_cohort(doby)) %>%    
  filter(is.na(nchartQuin)) %>%
  ggplot(aes(doby, avg_annual_tax_rate)) +
  geom_col() +
  geom_text(mapping = aes(label = scales::percent_format(accuracy = 1)(avg_annual_tax_rate)), 
            vjust = -1,
            position = position_dodge(width = 0.8)) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
  labs(x = "Birth Cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()

plotr(titler("Average tax rates for retirees will grow over time"),
      subtitler("Average annual tax rate at age 70"),
      chart13,
      caption,
      sourcer("DYNASIM4, ID963."),
      noter("Taxes include payroll taxes, federal and state income taxes, and Medicare surtaxes. Property, sales, and excise taxes are excluded.\nThe analysis assumes that current tax rules continue indefinitely and that scheduled Social Security benefits are paid."), 
      heights = c(2, 1, 30, 1, 1, 2))
```